Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-06 android-stacks.tex
threat
all over

       \item Aproximately 52\% of the reported stack traces were associated to programming mistackes. 
And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.

     \item  Cross-type exception wrappings (e.g., an OutOfMemoryError wrapped in a checked exception)
were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.

   \item Undocumented runtime exceptions signaled by third-party code (i.e., libraries and Android platform). 
    In this study, none of the third-party methods explicitly signaling runtime exceptions included such exceptions
    on ``exception specifications" (i.e., a list of exceptions that a method might raise during its execution).

@@

\begin{figure} \centering \includegraphics[scale=0.6]{stacktrace_bw.png}
\caption{Example of an exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

@@

\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled runtime exceptions related to programming
mistakes} as carefully as checked exceptions, specialy when developing
 library/framework reusable code. Doing so, the clients of such ibrary/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}.

@@

@@

To support this investigation, we developed a tool called ExceptionMiner (see Section ~\ref{sec:exceptionminer})
which extracts the exception stack traces embeeded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Such combination was needed to enable the identification of 
fragilities on the exception-related code.
Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.

@@

\begin{table}
  \centering
  \begin{tabular}{rp{29em}}
    \hline
    \bfseries{Signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} & If the exception is thrown on a method defined in Android Platform.\\
    \bfseries{app}     & If the exception is thrown on an application method.\\
    \bfseries{libcore} & If the exception is thrown on one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
    \bfseries{lib}     & If the exception is thrown on a method that was not defined by any of the elements above.\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

@@

Table ~\ref{tab:causes} summarized the results. It also illustrates the number of general Java exceptions - 
whose full name does not carry the reason for the exception to be thrown.
To ensure the quality of the analysis, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the resulting interrater agreement was high (96\%). We could also observe that
We can therefore assume that the exception names and Javadoc 
were enough for classifying the exception types analyzed.

@@

\begin{table}
  \centering
  \begin{tabular}{lr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
    \hline
      Programming Mistakes &  51,96\%\\ 
      Memory Restriction   &   4,10\% \\ 
      General Exceptions   &  6,70\%\\
    \hline
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Aproximately 52\% of the reported stack traces were associated to programming mistackes. 
And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
}
\end{minipage}}

@@

\emph{\textbf{Inspecting Exception Interfaces.}} To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
 signaled by  runtime environment (due to programming mistakes) since these exceptions 
should not be documented on the method signature.  Then we inspected the code each method 
(defined either on Android  Aplication Framework or on libraries) 
explicitly signaling a runtime exception not-related to programming mistakes.
As a result of this inspection, we could observe that only 1 method out of 118 inspected methods
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.
These ``undocumented runtime exceptions" represent a threat to system robustness, specially
when such exceptions are thrown then the third party code invoked inside the application 
(e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 Only 0,8\% of the methods documented the explicitly signaled runtime exceptions. 
}
\end{minipage}}

@@

This exceptions was thrown by the method getDeclaredMethods defined in java.lang.Class. 
The Java-side declaration of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.
 However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
violating the declaration. The Java compiler, could not detect this violation, because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usualy do not have access to the native implementations. 
Consequently, since it is not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).
%Among the Android projects on which this exception stack trace was reported we can cite:
%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618 stars).

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
An unexpected fragility was detected: a checked exception being thown and not specified  
on the ``exception interface" of the signaler methods.
}
\end{minipage}}

@@

Java is the only language that provides a hybrid exception model 
which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
runtime and checked). The best practices I and II are related to such type-related-behavior: 
(i) Errors should not be handled inside the system, since they usually represent unrecoverable conditions detected 
by the runtime environment (e.g.,  OutOfMemoryError); (ii) checked exceptions, on the other hand, should represent 
recoverable conditions.

Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
envolve different exception types (i.e., Error, Checked, Exception). We call 
such a wrapping a ``cross-type wrapping". Next we exemplify some of such
 wrappings and discuss why some of them may negatively impact the robustness.

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
on 46 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44\% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any instance of Throwable signaled 
in the context of an asynchronous task is converted in to a runtime and re-thrown.  Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
Such wrapping masks an unrecoverable error in a general exception - doing so it that can be mistakenly handled.

@@

\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}

@@

\textit{Empirical Studies analizing Java exception-related code.} 
Cabral and Marques~\cite{cabral2007exception} analyzed the
source-code of 32 open-source systems, both for Java and .NET. They
observed that the actions inside handlers were very simple (e.g., logging and present a
message to the user). Pingyu and Elbaum~\cite{Zhang12} 
analyzed the bug reports of 5 Android applications an observed that 29\% had to do with poor
exceptional handling code, this empirical study was used to motivate the development of a tool
aiming at amplifying existing tests to validate exception 
handling code associated with external resources. Our work mined the exception
stack traces embedded on issues reported 639 opensource Android projects
to identify common fragilities on the exception related code that can lead to 
failures such as uncaught exceptions.

@@

\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
approximately  50\% of such wrappings were performed on methods defined on Android
Platform. Taking a closer look on the methods responsible for such wrappings,
 we could observe that approximately 41\% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
(window) of an Android app.
Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
it is wrapped in a general Runtime exception and re-thrown.

@@

>>>>

       \item Approximately 52\% of the reported stack traces were associated to programming mistackes. 
And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.

       \item  4,4% of the reported exception stack traces ``undocumented runtime exceptions" signaled by third-party code. 
             (35 methods from Android platform and 44 from libraries).

     \item  Few cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception,
were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.

@@

\begin{figure} \centering \includegraphics[scale=0.5]{stacktrace_bw.png}
\caption{Example of an exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

@@

\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled runtime exceptions related to programming
mistakes} as carefully as checked exceptions, especially when developing
 library/framework reusable code. Doing so, the clients of such ibrary/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}.

@@

\begin{figure*} \centering \includegraphics[scale=0.3]{studyprocess2.png}
\caption{Study overview.} \label{fig:overview} \end{figure*}

@@

To support this investigation, we developed a tool called ExceptionMiner (see Section ~\ref{sec:exceptionminer})
which extracts the exception stack traces embeeded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Such combination was needed to enable the identification of 
fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.

@@

\begin{table}
  \centering
  \begin{tabular}{rp{29em}}
    \hline
    \bfseries{Signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} & If the exception is thrown on a method defined in Android Platform.\\
    \bfseries{app}     & If the exception is thrown on an method defined on an Android app.\\
    \bfseries{libcore} & If the exception is thrown on one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
    \bfseries{lib}     & If the exception is thrown on a method that was not defined by any of the elements above.\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

@@

Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
whose full name does not carry the reason for the exception to be thrown.
To ensure the quality of the analysis, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the resulting interrater agreement was high (96\%). We could also observe that
We can therefore assume that the exception names and Javadoc 
were enough for classifying the exception types analyzed.

@@

\begin{table}
  \centering
  \begin{tabular}{lr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
    \hline
      Programming Mistakes &  51,96\%\\ 
      %Memory Restriction   &   4,10\% \\ 
      General Exceptions   &  6,70\%\\
    \hline
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Approximately 52\% of the reported stack traces were associated to programming mistackes. 
And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
}
\end{minipage}}

@@

\emph{\textbf{Inspecting Exception Interfaces.}} To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
 signaled by  runtime environment (due to programming mistakes) since these exceptions 
should not be documented on the method signature.  Then we inspected the code each method 
(defined either on Android  Aplication Framework or on libraries) 
explicitly signaling a runtime exception not-related to programming mistakes.
Table~\ref{tab:runtimeinterface} present the result of this inspection. 
We could find 79 methods (both from libraries and Android Platform) that  explicitly threw a runtime exception 
without listing such exception on the exception interface (i.e., using 
throws clause on method signature). These methods were responsible for 
267 exception stack traces mined in this study.

This result is aligned with the results of other study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.
These ``undocumented runtime exceptions" represent a threat to system robustness, specially
when such exceptions are thrown then the third party code invoked inside the application 
(e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

Moreover, these stack traces could also pointed out another best practice violation: all android methods
 presented in Table~\ref{tab:runtimeinterface} were signaling direct instances of RuntimeException. 
Using such general exception, is considered a bad according to Java specification and compiled guidelines.

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform.
}
\end{minipage}}

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{lrr}
    \hline
 \bfseries{Origin}   & \bfseries{methods} &  \bfseries{stacks}  \\ 
\hline					
Libraries	& 	44	& 	205	\\					
\emph{android packages}   &  &   \\ 
android.app	& 	5	& 	16	\\
android.content	& 	2	& 	2	\\
android.database	& 	1	& 	1	\\
android.graphics	& 	5	& 	11	\\
android.hardware	& 	4	& 	5	\\
android.media	& 	3	& 	3	\\
android.opengl	& 	1	& 	8	\\
android.os	& 	11	& 	12	\\
android.text	& 	1	& 	1	\\
android.video	& 	1	& 	1	\\
android.providers	& 	1	& 	2	\\
\hline					
\bfseries{All android} 	&	35	&	62	\\
\hline					
\bfseries{All}	&	79	&	267	\\
\hline					
  \end{tabular}
\caption{Absence of exception interfaces on methods.}
\label{tab:runtimeinterface}
\end{table}

@@

This exceptions was thrown by the method getDeclaredMethods defined in java.lang.Class. 
The Java-side declaration of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.
 However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
violating the declaration. The Java compiler, could not detect this violation, because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usualy do not have access to the native implementations. 
Consequently, since it is not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).
%Among the Android projects on which this exception stack trace was reported we can cite:
%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618 stars).

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
An unexpected fragility was detected: a checked exception being thown and not specified  
on the ``exception interface" of the signaler methods.
Methods defined on 11 android packages signaled direct instances of RuntimeException.
}
\end{minipage}}

@@

Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
envolve different exception types (i.e., Error, Checked, Exception). We call 
such a wrapping a ``cross-type wrapping".

Java is the only language that provides a hybrid exception model 
which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
runtime and checked).

Next we exemplify some of such
 wrappings and discuss why some of them may negatively impact the robustness.

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
on 46 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44\% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet below show a general cach clause that convertes any Throwable in to a runtime and re-throw.  
Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
Such wrapping masks an unrecoverable error in a general exception - doing so it that can be mistakenly handled.

{\footnotesize
\begin{verbatim}
   try {
      ...
   } catch (InterruptedException e) {
      android.util.Log.w(..., e);
   } catch (ExecutionException e) {
      throw new RuntimeException("...",e.getCause());
   } catch (CancellationException e) {
      ...
   } catch (Throwable t) {
     throw new RuntimeException("...", t);
   }
\end{verbatim}
}

@@

%\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
%\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}

@@

\textit{Empirical Studies analizing Java exception-related code.} 
Cabral and Marques~\cite{cabral2007exception} analyzed the
source-code of 32 open-source systems, both for Java and .NET. They
observed that the actions inside handlers were very simple (e.g., logging and present a
message to the user). Pingyu and Elbaum~\cite{Zhang12} were the first to perform
an empirical investigation of issues (related to exception-related bugs) on Android projects.  
They perform a small scale study on which they manually inspected the issues of 
5 Android applications. They observed that 29\% had to do with poor
exceptional handling code, this empirical study was used to motivate the development of a tool
aiming at amplifying existing tests to validate exception 
handling code associated with external resources. This work inspired ours,
 which automaticaly mined the exception stack traces embedded on issues 
reported on 639 opensource Android projects. The goal of our study was
to identify common fragilities on the exception related code that can lead to 
failures such as uncaught exceptions.

@@

\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
approximately  50\% of such wrappings were performed on methods defined on Android
Platform. Taking a closer look on the methods responsible for such wrappings,
 we could observe that approximately 41\% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
(window) of an Android app.
The code snippet illustrates one the methods of 
ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
it is wrapped in a general Runtime exception and re-thrown.

@@

exemplos:

(java.lang.RuntimeException,android.hardware.Camera.native_autoFocus(Native Method)")"
